HealpixNet: Efficient spherical Convolutional Neural Network with Healpix sampling for cosmological applications

### 2018-07-27 scnn.tex
update figures
todos

\section{TODO list}
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
    \item (NP) Figures  A.15, 6, 5
    \item (TK) add small description of maps\_downsampled\_64.npz in the whole\_sphere notebook.
    \item (TK) update the README for data zip - file from NP
    \item (NP) Finalise readme on Github code
    \item (Waiting for Zenodo) (NP) Finalize the data on zenodo
    \item TK check part sphere notebook and whole sphere
    \item branch with notebooks with plots (results) NP
    \item Michael section 4.5 (efficient convolutions) and 3.2 (conv on graphs)
    \item NP move the repo to HealpixNet
	\item check if papers which appear as arxiv were not published in a conference or journal
    \item (TK) add info on comparison with histograms, so that it's just a demonstration
    \item (TK NP) if we have time make a new candidate for network plot (optional)
    \item (first version done) conclusions
    \item Should the heat example go to appendix?
    \item TK read the heat section again
    \item TK read the result section
    \item TK check the appendix
    \item TK optional - example of data (CMB (mask) + mass map + radio intensity)
    \item TK read everything
	\item MD add irregular (random) sampling to comparison with spherical harmonics? Could motivate people to think about different samplings.
	\item uniformize text size in figures
\end{itemize}

@@

In this work we present and release HealpixNet\footnote{\url{https://github.com/SwissDataScienceCenter/HealpixNet}}, an algorithm and software package which implements an efficient CNN on HEALPix maps. It is implemented with \pkg{TensorFlow} \citep{abadi2016tensorflow} and is intended to be easy to use out-of-the box for cosmological applications.
As HEALPix~\citep{gorski2005healpix} is the most popular sampling used in cosmology and astrophysics, we tailored our method to that particular sampling. However, only two elements depend on the sampling: (i) the choice of neighboring pixels when building the graph, and (ii) the choice of parent pixels when building the hierarchy.
It is also possible to use multiple HEALPix maps within our framework and use datasets in form of ``shells''. These shells can span the radial direction, so that a tomographic analysis can be performed, or different frequencies, in case of data from observations in radio frequencies. \nati{I do not know about that last point. Yes we can do it. But a) we did not do it and b) the filter will be spherical in all directions includeding the radial one... I am not sure this is a disirable features.}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{gaussian_filters_sphere}
	\caption{Comparison of convolution with the graph and the spherical harmonics ($N_{side} = 16$).
	Top: Diffusion of a unit of heat for different times $t$ using the graph.
	Bottom: spherical symmetric Gaussian smoothing for different $\sigma$ (arcmin).
	Relative difference between graph convolution and spherical smoothing: $10.4$\%, $5.7$\%, $4.8$\%, $3.8$\%.
	\todo{(i) Only three diffusion steps to have a bigger figure? \nati{Agreed! Will do soon!}}
\todo{(ii) The main issue with our filters is that their shape depends on the localization. It would be nice to measure the expectation and variance of the error over all locations. That would show how severe the distortion is.}
\nati{Let us keep that for the next paper. I am quite sure, it is negligible in this particular case.}}
	\label{fig:gaussian_filters_comparizon}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=\linewidth]{gaussian_filters_spectral}
\includegraphics[width=\linewidth]{gaussian_filters_gnomonic}
\includegraphics[width=\linewidth]{gaussian_filters_section}
\caption{Three different visualizations of the convolution kernel $K_t(x)=e^{-\tau t x}$.
Top: graph spectral domain.
Middle: gnomonic projection.
Bottom: section plot.}
\label{fig:gaussian_filters_visualization}
\end{figure*}

>>>>

\section{TODO list}
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
    \item (TK) add small description of maps\_downsampled\_64.npz in the whole\_sphere notebook.
    \item (TK) update the README for data zip - file from NP
    \item (NP) Finalise readme on Github code
    \item (Waiting for Zenodo) (NP) Finalize the data on zenodo
    \item TK check part sphere notebook and whole sphere
    \item branch with notebooks with plots (results) NP
    \item Michael section 4.5 (efficient convolutions) and 3.2 (conv on graphs)
    \item NP move the repo to HealpixNet
	\item check if papers which appear as arxiv were not published in a conference or journal
    \item (TK) add info on comparison with histograms, so that it's just a demonstration
    \item (TK NP) if we have time make a new candidate for network plot (optional)
    \item (first version done) conclusions
    \item Should the heat example go to appendix?
    \item TK read the heat section again
    \item TK read the result section
    \item TK check the appendix
    \item TK optional - example of data (CMB (mask) + mass map + radio intensity)
    \item TK read everything
	\item MD add irregular (random) sampling to comparison with spherical harmonics? Could motivate people to think about different samplings.
	\item uniformize text size in figures
\end{itemize}

@@

In this work we present and release HealpixNet\footnote{\url{https://github.com/SwissDataScienceCenter/HealpixNet}}, an algorithm and software package which implements an efficient CNN on HEALPix maps. It is implemented with \pkg{TensorFlow} \citep{abadi2016tensorflow} and is intended to be easy to use out-of-the box for cosmological applications.
As HEALPix~\citep{gorski2005healpix} is the most popular sampling used in cosmology and astrophysics, we tailored our method to that particular sampling. However, only two elements depend on the sampling: (i) the choice of neighboring pixels when building the graph, and (ii) the choice of parent pixels when building the hierarchy.
It is also possible to use multiple HEALPix maps within our framework and use datasets in form of ``shells''. These shells can span the radial direction, so that a tomographic analysis can be performed, or different frequencies, in case of data from observations in radio frequencies. \nati{I do not know about that last point. Yes we can do it. But a) we did not do it and b) the filter will be spherical in all directions includeding the radial one... I am not sure this is a disirable features.}
\todo{I think you got it wrong there. I guess the idea is that each shell would be a feature map, i.e., the graph signal would be multi-dimensional.}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{gaussian_filters_sphere}
	\caption{Comparison of convolution with the graph and the spherical harmonics ($N_{side} = 16$).
	Top: Diffusion of a unit of heat for different times $t$ using the graph.
	Bottom: spherical symmetric Gaussian smoothing for different $\sigma$ (arcmin).
	Relative difference between graph convolution and spherical smoothing: $10.4$\%, $4.8$\%, $3.8$\%.
\todo{The main issue with our filters is that their shape depends on the localization. It would be nice to measure the expectation and variance of the error over all locations. That would show how severe the distortion is.}
\nati{Let us keep that for the next paper. I am quite sure, it is negligible in this particular case.}}
	\label{fig:gaussian_filters_comparizon}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{gaussian_filters_spectral}
\includegraphics[width=\linewidth]{gaussian_filters_gnomonic}
\includegraphics[width=\linewidth]{gaussian_filters_section}
\caption{Three different visualizations of the convolution kernel $K_t(x)=e^{-\tau t x}$.
Top: graph spectral domain.
Middle: gnomonic projection.
Bottom: section plot.}
\label{fig:gaussian_filters_visualization}
\end{figure}
